### 2016-06-29 text/formal_background.tex
Move old sections. Divorce paper from definitions file.

Note that this section presents a collection of formal tools that are used in the subsequent sections of this paper where the contributions of this paper are presented. It is meant as a formal reference for the upcoming formal development. This section can be safely skipped or skimmed by the reader, who can return to these definitions punctually to understand the detailed formal underpinning of our approach.

\subsection{Typed Graphs}
\label{sec:typed_graphs}
% We begin this section by formally defining the graphs and graph instances used in our algorithm as fundamental structures.
% \begin{definition}{Typed Graph}
% \label{def:typed_graph}
% A typed graph is a triple $\langle V,E,\tau\rangle$ where $V$ is a finite set of
% vertices, $E\subseteq V\times V$ is a set of edges connecting the vertices and
% $\tau:V\rightarrow Type$ is a typing function for the elements of V, where $Type$ is
% a set of type names. Edges $(v,v')\in E$ are noted $v\rightarrow v'$. The set of
% all typed graphs is called $TG$.
% \end{definition}
% 
% Definition~\ref{def:typed_graph_homomorphism} describes when a typed graph is said to be homomorphic to another.
% 
% \begin{definition}{Typed Graph Homomorphism  / Typed Graph Instance}
% \label{def:typed_graph_homomorphism}
% 
% Let $\langle V,E,\tau\rangle=g,\langle V',E',\tau'\rangle=g'\in TG$ be typed
% graphs. There is a typed graph homomorphism between $g'$ and $g$ if and only if
% there exists a graph homomorphism $f:V'\rightarrow V$ such that for all $v'\in
% V$ we have that $\tau'(v')=\tau(f(v'))$. Note that, trivially, a typed graph
% homomorphism is a formal homomorphism. We denote an \emph{injective} typed graph
% homomorphism between typed graphs $g$ and $g'$ by $g' \vartriangleleft g$.
% \end{definition}
% 
% In the text that follows we will use the terms \emph{typed graph homomorphism}
% and \emph{typed graph instance} interchangeably.
% 
% Definition~\ref{def:typed_graph_strict_instance} is stricter than
% definition~\ref{def:typed_graph_homomorphism}, and indicates that there is a
% strict many-to-one type correspondence of the nodes of two graphs and their
% relations.
% 
% \begin{definition}{Typed Graph Strict Instance}
% \label{def:typed_graph_strict_instance}
% 
% Let $\langle V,E,\tau\rangle=g,\langle V',E',\tau\rangle=g'\in TG$ be typed
% graphs. $g'$ is a typed graph strict instance of $g$, noted $g'
% \blacktriangleleft g$, if and only if there exists a surjective
% typed graph homomorphism between $g'$ and $g$.
% 
% \end{definition}
% 
% In order to simplify the subsequent development we will often
% use the notation ${\langle V,E,\tau\rangle}^*$ to represent the transitive
% closure of the edges of typed graph  $\langle V,E,\tau\rangle$.

% \begin{definition}{Typed Graph}
% \label{def:typed_graph}
% A typed graph is a triple $\langle V,E,\tau\rangle$ where $V$ is a finite set of
% vertices, $E\subseteq V\times V$ is a set of edges connecting the vertices and
% $\tau:V\rightarrow Type$ is a typing function for the elements of V, where $Type$ is
% a set of type names. Edges $(v,v')\in E$ are noted $v\rightarrow v'$. The set of
% all typed graphs is called $TG$.
% \end{definition}

% \begin{definition}{Typed Graph and Indirect Typed Graph}
% \label{def:typed_graph}
% 
% A typed graph is a triple $\langle V,E,\tau\rangle$ where $V$ is a finite set of
% vertices, $E\subseteq V\times V$ is a finite set of directed edges connecting the vertices and $\tau:V\cup E \rightarrow Type\cup\{containment,reference\}$ is a typing function for the elements of V and E. Typing function $\tau$ is such that $\tau(v)\in Type$ if $v\in V$ and $\tau(e)\in \{containment,reference\}$\footnote{By using \emph{containment} and \emph{reference} as types for edges we allow modeling the different types of associations between the elements of a metamodel or a model. In particular, the fact that the subgraph of containment relations in a typed graph is acyclic models EMF containment associations.} if $e\in E$.  Edges $(v,v')\in E$ are noted $v\rightarrow v'$. We furthermore impose that the graph $\langle V,\{v\rightarrow v'\in E|\tau(v\rightarrow v')=containment\}\rangle$ is acyclic. The set of all typed graphs is called $TG$.
% 
% An indirect typed graph is a 4-tuple $\langle V,E,T,Il\rangle$, where $\langle V,E,T\rangle$ is a typed graph and $Il\subseteq E$ is a set of edges called \emph{indirect links}. The set of all indirect typed graphs is called $ITG$.
% \end{definition}

\CatchFileBetweenTags{\typedgraphtext}{text/definitions}{typedgraphtext}{\typedgraphtext}

\begin{definition}{Typed Graph\\}
\label{def:typed_graph}
\CatchFileBetweenTags{\typedgraph}{text/definitions}{typedgraph}{\typedgraph}
\end{definition}

\CatchFileBetweenTags{\typedgraphuniontext}{text/definitions}{typedgraphuniontext}{\typedgraphuniontext}

\begin{definition}{Typed Graph Union\\}
\label{def:typed_graph_union}
\CatchFileBetweenTags{\typedgraphunion}{text/definitions}{typedgraphunion}{\typedgraphunion}
\end{definition}

\CatchFileBetweenTags{\typedgraphhomomorphismtext}{text/definitions}{typedgraphhomomorphismtext}{\typedgraphhomomorphismtext}

\begin{definition}{Typed Graph Homomorphism\\}
\label{def:typed_graph_homomorphism}
\CatchFileBetweenTags{\typedgraphhomomorphism}{text/definitions}{typedgraphhomomorphism}{\typedgraphhomomorphism}
\end{definition}

Note that, trivially, a typed graph homomorphism is a graph homomorphism.

% \begin{lemma}{The Union of Two Typed Graph Homomorphisms is a Typed Graph Homomorphism}
% 
% Let $g=\langle V_g,E_g,st_g,\tau_g\rangle$, $g'=\langle
% V'_g,E'_g,st'_g,\tau'_g\rangle$, $h=\langle V_h,E_h,st_h,\tau_h\rangle$,
% $h'=\langle V'_h,E'_h,st'_h,\tau'_h\rangle$ be typed graphs, where $g$ and $h$
% are potentially joint, as well as $g'$ and $h'$. Let also $f:V_g\rightarrow V'_g$ and
% $i:V_h\rightarrow V'_h$ be a typed graph homomorphisms between $g$ and $g'$ and
% $h$ and $h'$, respectively. Let us show that $f\cup i$ is a typed graph homomorphism between $g\sqcup g'$ and $h\sqcup h'$.
% The proof is trivial if $g$ and $h$, as well as $h$ and $h'$ are disjoint. If $g$ and $h$ are joint over a vertex $v$, we know that $(f\cup i)(v)$
% \end{lemma}
% 
% \begin{lemma}{The Union of Two Surjective Typed Graph Homomorphisms is a Surjective Typed Graph Homomorphism}
% \levi{finish this proof}
% \end{lemma}

% \levi{changed the direction of the triangle for the surjective type graph homomorphism. Typed graph homomorphisms make sure that ALL connections between two nodes in the first graph also exists in the two mapped nodes of the second graph.}

% \begin{definition}{Strict Typed Graph Homomorphism}
% \label{def:strict_typed_graph_homomorphism}
% 
% Let $\langle V,E,\tau\rangle=g,\langle V',E',\tau\rangle=g'\in TG$ be typed
% graphs. $g'$ is a typed graph strict instance of $g$, noted $g'
% \blacktriangleleft g$, if and only if there exists a surjective
% typed graph homomorphism between $g'$ and $g$.

\CatchFileBetweenTags{\typedsubgraphtext}{text/definitions}{typedsubgraphtext}{\typedsubgraphtext}

\begin{definition}{Typed Subgraph\\}
\label{def:typedsubgraph}
\CatchFileBetweenTags{\typedsubgraph}{text/definitions}{typedsubgraph}{\typedsubgraph}

% An indirect typed graph $\langle V',E',\tau',Il\rangle \in ITG$ is an indirect typed
% subgraph of a typed graph $\langle V,E,\tau\rangle \in TG$, written $\langle
% V',E',\tau',Il\rangle \lhd \langle V,E,\tau\rangle$ iff:
% 
% \begin{enumerate}
% \item $\langle V',E'\setminus Il,\tau'\rangle \blacktriangleleft \langle V, E,\tau\rangle$
% \item if $v_i\rightarrow v_i'\in Il$ then there exists $v\rightarrow v'\in E_{c}^{*}$ where $\tau(v_i)=\tau(v)$, $\tau(v_i')=\tau(v')$ and $E_{c}^{*}$ is obtained by the transitive closure of $E_{c}=\{v\rightarrow v'\in E|\tau(v\rightarrow v')=containment\}$.
% \end{enumerate}
\end{definition}

\CatchFileBetweenTags{\typedgraphisomorphismtext}{text/definitions}{typedgraphisomorphismtext}{\typedgraphisomorphismtext}

\begin{definition}{Typed Graph Isomorphism\\}
\label{def:typed_graph_isomorphism}
\CatchFileBetweenTags{\typedgraphisomorphism}{text/definitions}{typedgraphisomorphism}{\typedgraphisomorphism}
\end{definition}
% \begin{definition}{Typed Graph Connected Component}
% \label{def:typed_graph_connected_component}
% 
% Let $g \in \textsc{TG}$ be a typed graph. We say that $\langle V,E,st,\tau\rangle \sqsubseteq g$ is a connected component of $g$ if for any pair of vertices $v,v'\in V$ there exists a path between $v$ and $v'$, where the direction of the edges between $v$ and $v'$ does not matter.
% \end{definition}

% \begin{definition}{Typed Graph Instance}
% \label{def:typed_graph_instance}
% 
% Let $\langle V,E,\tau\rangle=g,\langle V',E',\tau\rangle=g'\in TG$ be typed graphs.
% $g'$ is a typed graph instance of $g$, written $g'\Vdash g$, iff for all $v_{1}'\rightarrow v_{2}'\in E'$ there is a $v_{1}\rightarrow v_{2}\in E$ such that $\tau(v_{1}')=\tau(v_1)$, $\tau(v_{2}')=\tau(v_2)$ and $\tau(v_{1}'\rightarrow v_{2}')=\tau(v_{1}\rightarrow v_{2})$.
% \end{definition}

%Notice that we only enforce that connections between vertices of $g'$ must exist also in $g$ and have the same type.

% \levi{Note that each edge between two nodes in the first graph has to exist in the second graph and vice-versa.}
% \levi{in the sections that follow we will formally elaborate on relevant concepts that are introduced.}

\paragraph{\textbf{Notation:}}
\CatchFileBetweenTags{\notationformalbackground}{text/definitions}{notationformalbackground}{\notationformalbackground}

\subsection{Metamodel and Model-Related Constructs}
\label{sec:DSLTrans_formal}

% This section will formally describe the DSLTrans constructs that are required in
% our property-proving algorithm. The definitions presented strongly rely on the concepts of typed graphs, just introduced in \cref{sec:typed_graphs}.

We will start by introducing the notion of \emph{metamodel}, which in DSLTrans is used to type the input and output models of a DSLTrans transformation.  Two metamodels, the \emph{organization language} and \emph{gender language} are depicted in \cref{fig:squadmetamodel}.

\begin{definition}{Metamodel\\}
\label{def:metamodel}
\CatchFileBetweenTags{\metamodel}{text/definitions}{metamodel}{\metamodel}
\end{definition}

\CatchFileBetweenTags{\metamodeltext}{text/definitions}{metamodeltext}{\metamodeltext}

\begin{definition}{Expanded Metamodel\\}
\label{def:expanded_metamodel}
\CatchFileBetweenTags{\expandedmetamodel}{text/definitions}{expandedmetamodel}{\expandedmetamodel}
\end{definition}

\CatchFileBetweenTags{\expandedmetamodeltext}{text/definitions}{expandedmetamodeltext}{\expandedmetamodeltext}

\begin{definition}{Metamodel Instance\\}
\label{def:metamodelinstance}
\CatchFileBetweenTags{\metamodelinstance}{text/definitions}{metamodelinstance}{\metamodelinstance}
\end{definition}

\CatchFileBetweenTags{\metamodelinstancetext}{text/definitions}{metamodelinstancetext}{\metamodelinstancetext}

\begin{definition}{Containment Transitive Closure\\}
\label{def:instance_closure}
\CatchFileBetweenTags{\instanceclosure}{text/definitions}{instanceclosure}{\instanceclosure}

@@

\CatchFileBetweenTags{\instanceclosuretext}{text/definitions}{instanceclosuretext}{\instanceclosuretext}

\begin{definition}{Model\\}
\label{def:model}
\CatchFileBetweenTags{\model}{text/definitions}{model}{\model}
\end{definition}

\CatchFileBetweenTags{\modeltext}{text/definitions}{modeltext}{\modeltext}

\begin{definition}{Input-Output Model\\}
\label{def:input_output_model}
\CatchFileBetweenTags{\inputoutputmodel}{text/definitions}{inputoutputmodel}{\inputoutputmodel}
\end{definition}

%Change if needed
\CatchFileBetweenTags{\inputoutputmodeltext}{text/definitions}{inputoutputmodeltext}{\inputoutputmodeltext}

\begin{definition}{Metamodel Pattern and Indirect Metamodel Pattern\\}
\label{def:metamodel_pattern} 
\CatchFileBetweenTags{\metamodelpattern}{text/definitions}{metamodelpattern}{\metamodelpattern}
\end{definition}

\CatchFileBetweenTags{\metamodelpatterntext}{text/definitions}{metamodelpatterntext}{\metamodelpatterntext}

\subsection{Syntactic Transformation Constructs}
\label{sec:DSLTrans_syntax}

@@

\begin{definition}{Transformation Rule\\}
\label{def:transformation_rule}
\CatchFileBetweenTags{\transformationrule}{text/definitions}{transformationrule}{\transformationrule}
\end{definition}

\CatchFileBetweenTags{\transformationruletext}{text/definitions}{transformationruletext}{\transformationruletext}

\begin{definition}{Matcher of a Transformation Rule\\}
\label{def:back_match_transformation_rule}
\CatchFileBetweenTags{\backmatchtransformationrule}{text/definitions}{backmatchtransformationrule}{\backmatchtransformationrule}
\end{definition}

\Cref{def:back_match_transformation_rule} \CatchFileBetweenTags{\backmatchtransformationruletext}{text/definitions}{backmatchtransformationruletext}{\backmatchtransformationruletext}

\begin{definition}{Expanded Transformation Rule\\}
\label{def:transformation_rule_expansion}
\CatchFileBetweenTags{\expandedtransformationrule}{text/definitions}{expandedtransformationrule}{\expandedtransformationrule}
\end{definition}
\CatchFileBetweenTags{\expandedtransformationruletext}{text/definitions}{expandedtransformationruletext}{\expandedtransformationruletext}

\begin{definition}{Layer, Model Transformation\\}
\label{def:layer_transformation}
\CatchFileBetweenTags{\layertransformation}{text/definitions}{layertransformation}{\layertransformation}
\end{definition}

\CatchFileBetweenTags{\layertransformationtext}{text/definitions}{layertransformationtext}{\layertransformationtext}

\paragraph{\textbf{Notation:}}

\CatchFileBetweenTags{\notationtext}{text/definitions}{notationtext}{\notationtext}

% We naturally extend the notion of union (definition~\ref{def:typed_graph_union})
% to models (definition~\ref{def:metamodelmodel}), match-apply models
% (definition~\ref{def:match_apply_model}) and transformation rules 
% (definition~\ref{def:transformation_rule}). We also extend the notion of
% indirect typed subgraph (definition~\ref{def:indirecttypedsubgraph}) to
% transformation rules (definition~\ref{def:transformation_rule}) and match-apply
% models (definition~\ref{def:match_apply_model}). Finally, we extend the notion
% of typed graph equivalence (definition~\ref{def:typed_graph_equivalence}) to
% transformation rules (definition~\ref{def:transformation_rule}).

% \subsubsection{Transformation Language Semantics}
% 
% We will now address the semantics of the DSLTrans language. We will start by defining a match function that, given an input-output model and a transformation rule, returns all subgraphs of that input-output model where the rule's match pattern (including backward links) is found.  
% 
% % \levi{Need to be careful throughout all this section to understand whether we model the match and apply functions as functions or relations. Because of the non-deterministic generation of new vertices in the definition of apply function~\ref{def:apply_function}, relations are used in the definitions that follow to evaluate a transformation. This is an issue of the formalism used to model, should we make it explicit?}
% 
% \begin{definition}{Match Function}
% \label{def:match_function}
% 
% Let $m_{in}\in \textsc{Iom}^{sr}_{tr}$ be a input-output model and $rl\in \textsc{Rule}^{sr}_{tg}$ be a transformation
% rule. The $match : \textsc{Iom}^{sr}_{tg}\times \textsc{Rule}^{sr}_{tg}\rightarrow
% \mathcal{P}(\textsc{Iom}^{sr}_{tg})$ function is defined as follows: $$match_{rl}(m_{in})= \big\{ glue_{noind}\;|\;
% glue\sqsubseteq m_{in}^{*} \land glue \cong rl^{backM}_{trace}\big\}$$
% 
% where $glue$ is an input-output model. Additionally we have that: $rl_{trace}$ is a version of $rl$ where edges of type \emph{backward} have been replaced by edges of type \emph{trace}; $glue_{noind}$ is a version of $glue$ where the indirect links have been removed.
% % \levi{all backward links are replaced by trace links such that the matching using subgraph isomorphim can work.}
% % 
% % \levi{Where $rl^{-}$ is rule $rl$ without nodes in the apply part that are not connected to \emph{trace links}, as well as links adjacent to those nodes. $mg^{-ind}$ is the match graph without indirect links that come from the rule, given those cannot be part of the intermediate $MAM$ match result since that result is going to be united with the original $MAM$ (this was incorrect in the original SLE paper).}
% % \levi{The matcher function returns all graphs of the input MAM that are matched by a rule, including its backward links. Note that to find a subgraph ($\sqsubseteq$) of the input MAM that is isomorphic ($\cong$) to the rule being matched because the graph rewriting is achieved by joint union of the match graph extended by the new rewrite node, and the original MAM.}
% % \levi{TODO: review the $mg^{-ind}$ and the $rl^{-}$ notations.}
% 
% % Due to the fact that the $\cong$ relation is based on the notion of graph
% % isomorphism, permutations of the same match result may exist in the
% % $\big\{g\;|\; g\lhd m \land g \cong strip(tr)\big\}$ set. The --- undefined ---
% % $remove:\mathcal{P}(TR^{s}_{t})\rightarrow \mathcal{P}(TR^{s}_{t})$ function is such that
% % it removes such undesired permutations.
% 
% %The $strip:TR^{s}_{t}\rightarrow TR^{s}_{t}$ function is such that

>>>>

Note that this section presents a collection of formal tools that are used in the subsequent sections of this paper where the contributions of this paper are presented. It is meant as a formal reference for the upcoming formal development. %This section can be safely skipped or skimmed by the reader, who can return to these definitions punctually to understand the detailed formal underpinning of our approach.

\reviewer{There is a lot of redundancy in the paper, on several levels:
  - input-output-models, path conditions and properties are
    all defined very similarly. Couldn't you subsume them in a single
    definition and spell out the differences?}

\newcommand{\defineggprime}{Let $\langle V,E,(s,t),\tau,VT,ET\rangle = g$, and\\ $\langle V',E',(s',t'),\tau',VT',ET'\rangle = g'$, where $g, g' \in \textsc{TG}$.}

\newcommand{\ET}{\mathit{ET}}
\newcommand{\VT}{\mathit{VT}}

\bentley{Make sure that terminology is consistent between elements and vertices}

\subsection{Graph Structures and Functions}
\label{sec:typed_graphs}

\subsubsection*{Typed Graph}
We will start by introducing the notion of typed graph. A typed graph is the essential object we will use throughout our mathematical development. Typed graphs will be used to formalise all the important graph-like structures we will present in this paper. A typed graph is a directed multigraph (a graph allowing multiple edges between two vertices) where vertices and edges are typed.

\begin{definition}{Typed Graph\\}
\label{def:typed_graph}
A typed graph is a 6-tuple $\langle V,E,(s,t),\tau, \VT, \ET\rangle$ where:
\begin{itemize}
\item $V$ is a finite set of vertices
\item $E$ is a finite set of directed edges connecting the vertices $V$
\item $(s,t)$ is a pair of functions $s: E\rightarrow V$ and $t: E\rightarrow V$ that respectively provide the source and target vertices for each edge in the graph
\item Function $\tau:V\cup E\rightarrow \VT \cup \ET$ is a typing function for the elements of $V$ and $E$, where $\VT$ and $\ET$ are disjoint finite sets of vertex and edge type identifiers and $\tau(v)\in \VT$ if $v\in V$ and $\tau(e)\in \ET$ if $e\in E$
\item Edges $e\in E$ are noted $v\xrightarrow{e} v'$ if $s(e)=v$ and $t(e)=v'$, or simply e if the context is unambiguous
\item The set of all typed graphs is called $\textsc{TG}$
\item We define the empty graph to be a graph with $V = \emptyset$, $E = \emptyset$, and the other elements to be empty functions
\end{itemize}
\end{definition}

\subsubsection{Vertex and Edge Types}

Note that the our verification technique is geared toward model-driven engineering. Therefore, the types of vertices and edges in our typed graphs will be drawn from a particular \textit{metamodel}.

Sample metamodels for our running example can be seen in figures above \bentley{Fix.}. The particular metamodel for a typed graph will be indicated in the formalization when required, as (potentially) different metamodels will be present for the input and output graphs in DSLTrans rules.

We assume the existence of a number of functions in our formalization to aid the treatment of typing:

\begin{itemize}
\item $\mathit{isAbstract}: \mathit{VT} \rightarrow \{\mathit{true}, \mathit{false}\}$
\begin{itemize}
\item Where $\mathit{isAbstract}(\VT)$ returns $\mathit{true}$ iff $\VT$ is denoted as abstract (not able to be instantiated) by the metamodel, else $\mathit{false}$
\end{itemize}

\item $\mathit{isIndirect}: \mathit{ET} \rightarrow \{\mathit{true}, \mathit{false}\}$
\begin{itemize}
\item Where $\mathit{isIndirect}(\ET)$ returns $\mathit{true}$ iff $\ET$ is denoted as an \textit{indirect edge}, else $\mathit{false}$. The \textit{indirect} classification allows matching over indirect paths between vertices. This will be further clarified when required for our constructs.
\end{itemize}

\item $\mathit{matchesOver}: \{\VT \cup \ET\} \times \{\VT \cup \ET\} \rightarrow \{\mathit{true}, \mathit{false}\}$

\begin{itemize}
\item The purpose of this function is to assist in handling polymorphism in the metamodel
\item $\mathit{matchesOver}(T, T) \rightarrow \mathit{true}$
\item $\mathit{matchesOver}(T, T') \rightarrow \mathit{true}$ iff T' is a subclass of T in some defined partial ordering $\leq$ given by the metamodel
\item Otherwise, $\mathit{false}$
\end{itemize}
\end{itemize}

For simplification purposes, we will not represent edge cardinalities or containment relationships given by a metamodel in our notion of typed graph. Note that in fact we require these conditions to be relaxed to perform our graph rewriting.

\subsubsection*{Typed Subgraph}
We now define the useful notion of typed subgraph. As expected, a typed subgraph is simply a restriction of a typed graph to some of its vertices and edges.

\begin{definition}{Typed Subgraph\\}
\label{def:typedsubgraph}
\defineggprime

$g'$ is a typed subgraph of $g$, written $g'\sqsubseteq g$, iff $V'\subseteq V$, $E'\subseteq E$ and $\tau'=\tau |_{V'\cup E'}$.

@@

Given a typed graph $g\in \textsc{TG}$, will use the notation $Components(g)$ to describe the set of strongly connected typed graphs in $g$. Also, we will use the notation $g|_{t}$ to refer to the restriction of graph $g$ to its subgraph containing only edges of type $t$.

\subsubsection*{Typed Graph Union}
We now define how two typed graphs are united. A union of two typed graphs is trivially the set union of all the components of those two typed graphs. Note that we do not require the components of the two graphs to be disjoint, as in the following joint unions will be used to merge typed graphs.

\begin{definition}{Typed Graph Union\\}
\label{def:typed_graph_union}

The typed graph union is the function $\sqcup :\textsc{TG}\times \textsc{TG}\rightarrow \textsc{TG}$ defined as:
\begin{multline*}
\big\langle V,E,(s,t),\tau,\VT,\ET\big\rangle\;\sqcup\;\big\langle V',E',(s',t'),\tau',\VT',\ET'\big\rangle=\\
\big\langle V\cup V', E\cup E',(s\cup s', t\cup t'), \tau\cup \tau', \VT\cup \VT', \ET\cup ET'\big\rangle
\end{multline*}
\end{definition}

Note: as a reviewer helpfully pointed out, we require that $s \cup s'$, $t \cup t'$, and $\tau \cup \tau'$ coincide on common elements. However, this can be assumed w. l. o. g..

\bentley{Do we need typed graph intersection? Or to mention what it means for two graphs to be disjoint?}

\subsubsection*{Typed Graph Homomorphism}
For the formal development of our technique, we are interested in relations between typed graphs that are structure-preserving, i.e. homomorphisms. Homomorphisms between typed graphs preserve not only structure, but also the types of vertices and edges that are mapped. Note that, trivially, a typed graph homomorphism is a graph homomorphism.

\begin{definition}{Typed Graph Homomorphism\\}
\label{def:typed_graph_homomorphism}
\defineggprime

A typed graph homomorphism between $g$ and $g'$ is a function $f: f_v \cup f_e$ such that:
\begin{itemize}
\item $f_v: V\rightarrow V'$
\item $f_e: E\rightarrow E'$
\item $\forall v_1 \xrightarrow{e} v_2\in E$:
\begin{itemize}
\item Let $ v_1' = f_v(v_1), v_2' = f_v(v_2), e' = f_e(e)$
\item $v_1', v_2' \in V', e' \in E'$
\item $s'(e') = v_1'$, $t'(e') = v_2'$
\item $\mathit{matchesOver}(\tau(v_1), \tau(v_1') \land \\\mathit{matchesOver}(\tau(v_2), \tau(v_2') \land\\ \mathit{matchesOver}(\tau(e), \tau(e')$
\end{itemize}

%, where $\tau(v_1)=\tau'(f(v_1))$, $\tau(v_2)=\tau'(f(v_2))$ and also $\tau(e)=\tau(e')$.
\item The domain of $f$ is noted $Dom(f)$ and the co-domain of $f$ is noted $CoDom(f)$.
\end{itemize}  
\end{definition}

\bentley{Maybe put `inj' and `sub' in the homomorphism to reduce number of symbols}
 When an \emph{injective} typed graph homomorphism $f$ exists between $g$ and $g'$ we write $g \stackrel{f}{\vartriangleleft} g'$, or simply $g \vartriangleleft g'$ when the context is unambiguous. When a \emph{surjective} typed graph homomorphism $f$ exists between typed graphs $g$ and $g'$ we write $g \stackrel{f}{\blacktriangleleft} g'$, or also simply $g \blacktriangleleft g'$ in an unambiguous context.

\subsubsection*{Typed Graph Edge Homomorphism}

Our technique also requires a form of graph homomorphism that primarily focuses on matching the edges of the graphs. This homomorphism will be an injective match regarding the edges, but note that a particular vertex in the source graph may match onto multiple vertices in the target graph. \bentley{What's the term for this?}

\begin{definition}{Typed Graph Edge Homomorphism\\}
\label{def:typed_graph_edge_homomorphism}
\defineggprime

A typed graph edge homomorphism between $g$ and $g'$ is a function $h: h_v \cup h_e$ such that:
\begin{itemize}
\item $h_v: V'\rightarrow V$
\begin{itemize}
\item Note that $h_v$ is an function from a vertex in the target graph onto a vertex in the source graph
\end{itemize}
\item $h_e: E\rightarrow E'$
\item $\forall v_1 \xrightarrow{e} v_2\in E$:
\begin{itemize}
\item Let $e' = h_e(e), v_1' = s'(e'), v_2' = t'(e')$
\item $v_1', v_2' \in V', e' \in E'$
\item $v_1 = h_v(v_1'), v_2 = h_v(v_2')$
\item $\mathit{matchesOver}(\tau(v_1), \tau(v_1') \land \\\mathit{matchesOver}(\tau(v_2), \tau(v_2') \land\\ \mathit{matchesOver}(\tau(e), \tau(e')$
\end{itemize}

\end{itemize}  
\end{definition}

Again, we wish to highlight the fact that this homomorphism is different than Definition~\ref{def:typed_graph_homomorphism} in the vertex matching function. A vertex in the target graph matches onto a vertex in the source graph, with multiple target vertices matching onto the same source vertex. 
\bentley{Create diagram to illustrate}

This unusual homomorphism allows our technique to `split' our source graph over multiple locations in the target graph. This is critical, as our target graphs are constructed in a way that allows for vertex duplication. That is, two vertices of a particular type in the target graph may only represent one underlying element, as discussed in Section~\ref{sec:abstraction_relation}. Constructing this morphism allows our prover implementation to avoid explicitly creating `disambiguated' graphs where these duplications are resolved.

\subsection*{Typed Graph Isomorphism}
Two typed graphs are said to be isomorphic if they have exactly the same shape and related vertices and edges have the same type.

\begin{definition}{Typed Graph Isomorphism\\}
\label{def:typed_graph_isomorphism}
\defineggprime

We say that $g$ and $g'$ are isomorphic, written $g\cong g'$, iff there exists a bijective typed graph homomorphism $f:V\rightarrow V'$ such that $f^{-1}:V'\rightarrow V$ is a typed graph homomorphism.
\end{definition}

\subsubsection*{Transitive Closure}

DSLTrans matching constructs allow for the matching of indirect links, where as described in Section~\ref{subsec:DSLTrans_constructs}, rules may match over indirect paths between elements. This transitive closure allow for the explicit creation of all edges implied by these indirect links to aid in matching.

\begin{definition}{Transitive Closure\\}
\label{def:instance_closure}

Let $g$ be a graph $\langle V,E,(s, t),\tau\rangle \in \textsc{TG}$. Then the transitive closure $g^{*} = \langle V,E',(s, t),\tau\rangle \in \textsc{TG}$, where:

$E' = E \cup $ the transitive closure of $\big\{v\xrightarrow{e}v' | \mathit{isIndirect(e)} \big\}$
\end{definition}

Given a graph, its transitive closure includes, besides the original graph, all the edges belonging to the transitive closure of indirect links in that graph. Note that these transitive edges are given by the function $\mathit{isIndirect}$.

\bentley{I believe that the indirect edges don't need to be removed. There's a bit of handwaving, so maybe there are edge duplicates?}

In the definitions that follow we will use the $*$ notation, as in \cref{def:instance_closure}, to denote the transitive closure of our structures.

\subsection{DSLTrans Constructs}
\label{sec:DSLTrans_formal}

@@

As discussed in Section~\ref{subsec:DSLTrans_syntax}, DSLTrans transformations are composed of rules arranged in layers.

\subsubsection*{DSLTrans Transformation Rule}

A transformation rule is the elemental block of a DSLTrans transformation. Several transformation rules can be observed in the Police Station transformation in \cref{fig:dsltransformation}.

A transformation rule includes a non-empty match pattern and a non-empty apply pattern. This is also known in the model transformation literature as a rule's \emph{left hand side} and \emph{right hand side}. A match pattern can include indirect links that are used to transitively match containment relations in a model. The apply pattern of a rule always contains at least one apply element that is not connected to a backward link or an edge, meaning in practice that a rule will always produce something and not only match. An apply pattern does not include indirect links as it is used only for the construction of parts of instances of a metamodel.

A rule may also contain a negative application condition (NAC) which if matched over the source graph prevents application of the rule.

A transformation rule also includes backward links, as informally introduced in \cref{subsec:DSLTrans_constructs}. As described in that section, backward links define dependencies between rules. \bentley{Expand}

\begin{definition}{DSLTrans Transformation Rule\\}
\label{def:transformation_rule}

A DSLTrans transformation rule is a four-tuple $\big\langle \mathit{NAC}$, $\mathit{Match}$, $\mathit{Apply}$, $\mathit{backward}\big\rangle$, where:

\begin{itemize}
\item $\mathit{NAC}$, $\mathit{Match}, \mathit{Apply} \in \textsc{TG}$
\item $\mathit{Match}$ and $\mathit{Apply} $ are non-empty and are disjoint
\item $\mathit{NAC}$ may be an empty graph, and is disjoint from both $\mathit{Match}$ and $\mathit{Apply} $
\item $\mathit{backward} = \{E_{back}, (s_{back}, t_{back})\}$
\end{itemize}

Note that when we require an element of the $\mathit{Match}$, $\mathit{Apply}$, or $\mathit{NAC}$ graphs, such as the vertices, we will index the required element. For example, the vertices for the \textit{Match} graph will be $V_{\textit{Match}}$.

\begin{itemize}
\item $E_{back}$ contains the backward links
\begin{itemize}
\item $E_{back} $ is disjoint from $E_{Match}$,  $E_{Apply}$, and $E_{NAC}$
\end{itemize}
\item $(s_{back}, t_{back})$ is a pair of functions $s_{back}: E_{back}\rightarrow V_{\textit{Apply}}$ and $t_{back}: E_{back}\rightarrow V_{\textit{Match}}$ that respectively provide the source and target vertices for each backward link
\begin{itemize}
\item Note the source of backward links is a vertex in $V_{\textit{Match}}$ while the target is a vertex in $V_{\textit{Apply}}$
\end{itemize}
\end{itemize}

\textsc{Rules} is the set of all rules.

We additionally impose that for the well-formedness of a DSLTrans rule, there always exists an element or edge to be created in the \textit{Apply} graph to be created. That is, either there are edges to be created in the \textit{Apply} graph, or there is a vertex in the \textit{Apply} graph that is not the source of a backward link.

\begin{itemize}
\item $E_{\textit{Apply}} \neq \emptyset \lor$
\item $ \exists v \in V_{\textit{Apply}} | \big\{\forall e \in E_{back}: s_{back}(e) \neq v \big\}$
\end{itemize}

\subsubsection*{DSLTrans Layer and Transformation}

\cref{def:layer} and \cref{def:transformation} formalise the abstract syntax of a model transformation, introduced in \cref{sec:dsltrans}. An example of a model transformation can be observed in \cref{fig:dsltransformation}, the Police Station transformation. As expected, a DSLTrans transformation is composed of a sequence of layers where each layer is composed of a set of rules.

\begin{definition}{Layer\\}
\label{def:layer}
A layer is a finite set of transformation rules: $\{r_0, r_1, \dots, r_n | r_i \in \textsc{Rules}\}$.

The set of all layers is denoted $\textsc{Layers}$. 
\end{definition}

Note that the order of the rules within the layer does not matter, due to the semantics of DSLTrans rule execution.

\begin{definition}{DSLTrans Transformation\\}
\label{def:transformation}

A DSLTrans transformation is a finite list of layers denoted $\{l_0, l_1, \dots, l_n | l \in \textsc{Layers}\}$. Note that the order of layers in a transformation is important.

The set of all transformations is denoted $\textsc{Transforms}$.

\subsubsection*{Input-Output Model}

To describe the semantics of a DSLTrans model transformation, we must define an \textit{input-output model} construct. This input-output model allows the representation of intermediate operational states during the execution of a model transformation.

The structure of a input-output model is intentionally very similar to a DSLTrans rule. There is one typed graph representing the input model and another typed graph representing the output model. As well, the construct contains a set of edges, named \emph{traceability links}, for keeping a history of which elements in the output model originated from which elements in the input model.

\begin{definition}{Input-Output Model\\}
\label{def:input_output_model}

An input-output model rule is a three-tuple $\big\langle \mathit{Input}$, $\mathit{Output}$, $\mathit{trace}\big\rangle$, where:
